### 2025-07-24 conteudo/tool/architecture.tex
another grammar typos fix

The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
The flow of how the Preprocessor works is as follow: Function Parser receives the codebase the user is interested in,
extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
The Code Duplication Finder iterates over every pair of files in the Temporary Codebase, checks if they are code clones, and, 
if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple 
\textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates 
of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized in the Code Duplication Finder. ,

@@

Regarding the programming language's function extractor, we approached this problem so that anyone comfortable with a language can adapt our approach to their specific programming language. Usually, code blocks can be easily extracted from source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to infer if they represent a function's body given their depth in the code context and a few lines before the start of the code block. We followed this approach to implement the function extractor for the supported languages. As an alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to extract the functions, which is a more complex task than our approach. For the programming languages that we support, we implemented them as follows:

@@

For a source file in the C programming language, we extract the code blocks with depth 0; 
as per the programming language’s grammar, functions are defined in a global context at depth 0. 
To extract the code blocks, we first need to identify the portions of code that are comments or 
\textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
the curly brackets in depth 0, and with that, we find the body of the functions.

To extract the header of the functions, we start at the open curly brackets and go back in the 
source file considering the programming language grammar and the portions of code that have been 
marked as irrelevant because of comments and \textit{\#define} directives.
There are other elements in depth zero that are not functions. We analyze whether the code blocks 
are functions at the same time that we extract the header of the code block.

@@

For a source file in Java, we extract the code blocks with depth one and some code lines before 
defining them. Then, we check if the code blocks are functions by analyzing whether the first 
non-empty character before the curly brackets of the code block is a closing parenthesis.

@@

The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing 
functions from the input codebase. For each pair of files, we execute a code duplication detection method that computes 
a metric measuring how similar the pair of files is, which we refer to as similarity. 
If the similarity is greater than or equal to the minimum similarity threshold 
(a parameter provided by the user during the Preprocessor), we store this pair of functions along with its 
similarity in the Code Duplication Database.

We implemented two methods for code duplication detection in the tool, which the user can 
choose to use. The first method is based on text similarity, and the second is simpler and 
based on the number of equal lines. The experiments and tests in this research were done 
using the text similarity method. The purpose of the second method is to demonstrate that 
the tool can have multiple duplication detection methods.

@@

This method was chosen for its claimed performance, 
programming language independence, and the fact that it does not require 
compilable code, which is expected in the temporary codebase, as it does 
not contain complete code artifacts. We expect our method to yield good 
results for Type-1 and Type-2 code duplication types while performing 
poorly for the other types. It is possible to change the code duplication 
detection method to one of the state-of-the-art methods if needed.

@@

This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. To compute the number of equal lines, we use
the \textit{diff} command built-in in the Linux environment \citep{diffcommand}.

@@

The Query Responder step is the component the user executes multiple times to consult information 
about duplicated functions in a codebase processed by the Preprocessor. 
It utilizes the Temporary Codebase and the Code Duplication Database as inputs to answer user requests, 
such as querying for duplicated functions above a certain similarity threshold or provide counts of such functions.
This component was developed to be extensible, allowing other query functionalities to be easily added. 
The specific functionalities provided to the user by the Query Responder, which include the 
Duplication Explorer, Function Information, and Duplication Report, are described in Section \ref{subsec:func}.

>>>>

The Preprocessor contains two main components, the Function Breaker and the Code Duplication Finder. 
The flow of how the Preprocessor works is as follows: Function Parser receives the codebase the user is interested in,
extracts the functions of the codebase along with metadata, and creates a new temporary codebase where the functions extracted become new code files. 
The Code Duplication Finder iterates over every pair of files in the Temporary Codebase, checks if they are code clones, and, 
if so, saves them in the Code Duplication Database, which is a text file that stores every code duplication as a triple 
\textbf{<function1,function2,similarity>}, where \textbf{function1} and \textbf{function2} are the functions that are duplicates 
of each other, and \textbf{similarity} is the metric given by the code duplication detection method utilized in the Code Duplication Finder. ,

@@

Regarding the programming language's function extractor, we approached this problem so that anyone 
comfortable with a language can adapt our approach to their specific programming language. 
Usually, code blocks can be easily extracted from source code files, independently of whether the 
code blocks are defined by curly brackets or by indentation, and it is possible to infer if 
they represent a function's body given their depth in the code context and a few lines before the 
start of the code block. We followed this approach to implement the function extractor for the 
supported languages. As an alternative approach, we can build the syntax tree \citep{compiler} 
of the programming language and use it to extract the functions, which is a more complex task than 
our approach. For the programming languages that we support, we implemented them as follows:

@@

For a source file in the C programming language, we extract the code blocks with depth 0. 
As per the programming language’s grammar, functions are defined in a global context at depth 0. 
To extract the code blocks, we first need to identify the portions of code that are comments or 
\textit{\#define} directives and mark them as irrelevant. After that, it is simple work to find 
the curly brackets in depth 0; with that, we find the body of the functions.

To extract the header of the functions, we start at the open curly brackets and go back in the 
source file, considering the programming language grammar and the portions of code 
marked as irrelevant because of comments and \textit{\#define} directives.
Other elements in depth zero are not functions; we analyze whether the code blocks 
are functions at the same time that we extract the header of the code block.

@@

We extract the code blocks with depth one and some code lines for a source file in Java before 
defining them. Then, we check if the code blocks are functions by analyzing whether the first 
non-empty character before the curly brackets of the code block is a closing parenthesis.

@@

The Code Duplication Finder iterates through every pair of source code files in the temporary codebase, representing 
functions from the input codebase. For each pair of files, we execute a code duplication detection method that computes 
a metric measuring how similar the pair of files is, which we refer to as similarity. 
If the similarity is greater than or equal to the minimum similarity threshold 
(a parameter provided by the user during the Preprocessor). We store this pair of functions and its 
similarity in the Code Duplication Database.

We implemented two methods for code duplication detection in the tool, which the user can 
use. The first method is based on text similarity, and the second is simpler and 
based on the number of equal lines. The experiments and tests in this research were done 
using the text similarity method. The purpose of the second method is to demonstrate that 
the tool can have multiple duplication detection methods.

@@

We chose this method for its claimed performance, 
programming language independence, and the fact that it does not require 
compilable code, which is expected in the temporary codebase, as it does 
not contain complete code artifacts. We expect our method to yield good 
results for Type-1 and Type-2 code duplication types while performing 
poorly for others. Changing the code duplication
detection method to one of the state-of-the-art methods is possible if needed.

@@

This code duplication method is considerably simpler and more explainable than the text similarity method, 
as the metric is the ratio of common lines between the two functions. 
We use the \textit{diff} command built-in the Linux environment \citep{diffcommand}
to compute the number of equal lines.

@@

The Query Responder step is the component that the user executes multiple times to consult information 
about duplicated functions in a codebase processed by the Preprocessor. 
It utilizes the Temporary Codebase and the Code Duplication Database as inputs to answer user requests, 
such as querying for duplicated functions above a certain similarity threshold or providing counts of such functions.
This component was developed to be extensible, allowing other query functionalities to be easily added. 
The specific functionalities provided to the user by the Query Responder, which include the 
Duplication Explorer, Function Information, and Duplication Report, are described in Section \ref{subsec:func}.
